Databrary: A digital data library for sharing research video

### 2015-01-16 jcdl-15/paper.tex
sending up some notes and edits
probably have more, but overall I think now is a good time, since we have some structure and details, to reflect on the overall gist of what we are trying to communicate, that is to keep site of the larger picture while discussing the nitty gritty.

\begin{abstract}

Video captures the complexity, richness, and diversity of behavior
unlike any other tool. As a result, large numbers of researchers who
study human or animal behavior employ video. Video documents itself to a
large degree, and therefore has significant potential for data re-use.
Still, researchers rarely share video. Video often contains information
about personal identities, so considerations about research ethics pose
challenges to sharing. The relatively large size of video files and
diversity of formats pose technical challenges. In this paper, we will
describe how the web-based Databrary data library has overcome the most
significant barriers to sharing video within the developmental and
learning sciences community, including solutions to maintaining
participant privacy, data tagging, and data management.
\end{abstract}

@@

Video plays a key role in many areas of human research, particularly in developmental psychology, education, and is even a primary data source ...
Unfortunately, in standard practice, due largely to the perceived requirements about privacy restrictions, the use of these video data is isolated to the researchers who collected them.
The goal of the Databrary project is to establish policies and tools to allow video data to be documented, organized, and shared so that other researchers can discover and re-purpose existing videos for their own research interest.
Accordingly, Databrary, with developmental psychology as a use case, has created a web-based library and organizational structures to accommodate existing video data, capture contextual information necessary for re-use, and standardize across datasets without placing undue burden on data contributors.

% this is questionable...
Most existing research data repositories provide either unstructured storage of arbitrary files, or have rigorously structured schema that require contributors to properly extract and format contributions.
While the former is well suited for archival of entire datasets and general-purpose hosting of materials, it does not facilitate discoverability, understanding, or re-use across various datasets.
The latter, on the other hand, generate large-scale, homogonous data sources for meta- and re-analysis, but limited to specific domains and research questions, losing the full richness and innovation of the source materials and field.
We have attempted here to adopt a middle ground: the bar for contributing a collection, but when data are provided we encourage a good deal of common structure, either by capturing data as they are collected, or by curation after the fact.
It is this structure that is the focus of this paper.
% we're a hybrid?

We have intentionally chosen to limit the bulk of the data hosted in the library to human-consumable, self-documenting, timeseries data: video and audio.
These data can and largely are consumed and analyzed, at least initially, without sophisticated analysis tools: once digitally decoded, humans can glean the bulk of the richness of such data directly.
Similarly, these data can often be used on their own or with minimal explanation, as the nature of the activity and participants are readily apparent.
This naturally aids in the problems of documentation and re-use that plague most flat-file or dense datasets.
On the other hand, since automated analysis of videos is an active and immature research area, the problems of search, discoverability, and meta-analysis remain.
We attempt to address these issues by augmenting the primary data with additional, structured annotations and metadata.

@@

@@

Researchers record activities with individuals either for the purpose of a later, closer review or analysis, or simply to document and verify measurements and observations made during the collection.
Thus, the video itself is often the core item of research, representing a single data collection, experiment, trial, visit, or case depending on the nature of the research.
However, in many cases a single session can comprise multiple videos: from serial data streams due to recording equipment being reset or producing multiple files, and/or in parallel from multiple data capture sources that overlap in time.
These primary recordings may also be augmented with other forms of timeseries data such as motion capture, eye tracking, or physiological recordings, all of which have specific and critical temporal relationships with the primary data.

@@

Once the timeframe of a session is established, this can allow temporally-positioned annotations to be attached to the session.
% this is sort of an aside:
In fact, this practice is very common in research when analyzing recordings, and specialized desktop software tools are employed to review video and audio and annotate events, utterances, actions, or phases at any temporal scale, from individual samples/frames to the entire session.
Most of these tools currently focus on a single recording file, however, and despite the plethora of tools and features, there is little interoperability.

We currently allow keyword-based tagging and comment-based discussions on any temporal segment of a session.
These annotations could be applied manually by users or by automated video analysis systems.
We expect that these annotations will provide an index target for additional discovery capabilities for users looking for particular features within data.

>>>>

\begin{abstract}
%%dg being that the conference them is big data, should we be trying to be more explicit about that?
% big data is generally referring to "more rows" rather than larger files. I think the closest we get is the size of video files
% which is already noted below. Just wondering if it makes sense to be more explicit or not.
Video captures the complexity, richness, and diversity of behavior
unlike any other tool. As a result, large numbers of researchers who
study human or animal behavior employ video. Video documents itself to a
large degree, and therefore has significant potential for data re-use.
Still, researchers rarely share video. Video often contains information
about personal identities, so considerations about research ethics pose
challenges to sharing. The relatively large size of video files and
diversity of formats pose technical challenges. In this paper, we will
describe how the web-based Databrary data library has overcome the most
significant barriers to sharing video within the developmental and
learning sciences community, including solutions to maintaining
participant privacy, data tagging, and data management.
%%dg also preservation and discoverability - though maybe that's included in data management
\end{abstract}

@@

Video plays a key role in many areas of human research, particularly in developmental psychology, education, and is even a primary data source ...
%%dg unclear what is meant by primary data source in this context
Unfortunately, in standard practice, the use of these video data is isolated to the researchers who collected them because of the perceived requirements about privacy restrictions.
The goal of the Databrary project is to establish policies and tools to allow video data to be documented, organized, and shared so that other researchers can discover and re-purpose existing videos for their own research interest.
Accordingly, Databrary, with a focus on developmental psychology research, has created a web-based library to accommodate the storage and organization of video data, capture contextual information necessary for re-use, and standardize across datasets without placing undue burden on data contributors.

% this is questionable...
Most existing research data repositories provide either unstructured storage of arbitrary files, or have rigorously structured schema that require contributors to properly extract and format contributions.
%%dg should pick examples to illustrate both sides of the spectrum (ICPSR - ???)
While the former is well suited for archival of entire datasets and general-purpose hosting of materials, it does not facilitate discoverability, understanding, or re-use across various datasets.
%%dg i'd say that the general purpose ones are fine on discoverability, but they might no facilitate re-use or reflect very much of the original 
%context of the data
The latter, on the other hand, generate large-scale, homogonous data sources for meta- and re-analysis, but limited to specific domains and research questions, losing the full richness and innovation of the source materials and field.
Databrary attempts to strike a balance by creating a common structure that is general enough to provide the flexibility for incorporating heterogenous datasets while at the same time facilitating a depth of access to configure data more to the original context in which it was collected and analyzed. While the architecture of the backend allows us to curate and include diverse datasets, another important component is a well-define user experience that allows researchers to define how their data is kept in Databrary by uploading it as they collect it. 
The following provides further detail on the design and decision-making of this structure.
% we're a hybrid?
%%dg more like a system informed by the pitfalls of committing solely to either extreme.

%%dg perhaps better to frame below as the affordances and challenges of working with video data. 
%That is..."A key component to understanding the design decisions made in creating Databrary is the unique affordances and challenges of video and audio data"
%Affordances
We have intentionally chosen to limit the bulk of the data hosted in the library to human-consumable, self-documenting, timeseries data: video and audio.
These data can and largely are consumed and analyzed, at least initially, without sophisticated analysis tools: once digitally decoded, humans can glean the bulk of the richness of such data directly.
Similarly, these data can often be used on their own or with minimal explanation, as the nature of the activity and participants are readily apparent.
This naturally aids in the problems of documentation and re-use that plague most flat-file or dense datasets.
%Challenges
On the other hand, since automated analysis of videos is an active yet immature research area, the problems of search, discoverability, and meta-analysis over these media remain.
We attempt to address these issues by augmenting the primary video data with additional, structured annotations and metadata.
%%dg probably need to make this section more cohesive, but the main gist is 1) We focus on video data, 2)video data has these inherent challenges. 3) At the same time there are already challenges involved in creating a research data repository, independant of the format of the data travels on. 4) This paper dicusses the challenges we have resolved and the challenges we have coming up.

@@

%%dg would a figure illustration be helpful here? do we already have one?

@@

(Behvioral|Developmental) Science Researchers record activities with individuals either for the purpose of a later, closer review or analysis, or simply to document and verify measurements and observations made during the collection.
Thus, the video itself is often the core item of research, representing a single data collection, experiment, trial, visit, or case depending on the nature of the research.
However, in many cases a single session can comprise multiple videos: from serial data streams due to recording equipment being reset or producing multiple files, and/or in parallel from multiple data capture sources that overlap in time.
These primary recordings may also be augmented with other forms of timeseries data such as motion capture, eye tracking, or physiological recordings, all of which have specific and critical temporal relationships with the primary data.

@@

Once the timeframe of a session is established, this can allow temporally-positioned annotations to be attached to the session.
% this is sort of an aside:
%%dg maybe the value here is that we've designed things with this specific type of research in mind.
In fact, this practice is very common in research when analyzing recordings, and specialized desktop software tools are employed to review video and audio and annotate events, utterances, actions, or phases at any temporal scale, from individual samples/frames to the entire session.
Most of these tools currently focus on a single recording file, however, and despite the plethora of tools and features, there is little interoperability.

In addition to providing an organizational structure that allows users to visualize data in ways familiar to their research, we also currently allow keyword-based tagging and comment-based discussions on any temporal segment of a session.
These annotations could be applied manually by users or by automated video analysis systems.
We expect that these annotations will provide an index target for additional discovery capabilities for users looking for particular features within data.

%%dg, i guess it's crucial to make sure the audience understands that one of the specific challenges for us is that we're not just storing comprehensive single video files, but multiple files that capture the same phenomenon and how do we keep all those together in a meaningful way...to differentiate us from being "just another you tube" .. which i guess is what is already being said above. Though might need to do a little more hand holding for the reader as to why this is an issue in general and why it's important to even those who might be faced with different types of video data to collect (? maybe I'm rambling now ?).
